---
title: "p8105_hw6_mk4022"
author: "Mary Knoop"
date: "11/23/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(modelr)
library(knitr)
set.seed(1)

```


**Question 1:**   
*Part A:*  Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).
```{r}
bw_data=
  read_csv ("./data/birthweight.csv") %>% 
  janitor::clean_names() %>%
  mutate(babysex = factor(babysex, levels =  c("1", "2"), labels = c("male", "female")),
         frace = factor(frace, levels =  c("1", "2", "3", "4", "8", "9"), labels = c("white", "black", "asian", "puerto rican", "other", "unknown")),
         malform = factor(malform, levels =  c("0", "1"), labels = c("absent", "present")),
         mrace = factor(mrace, levels =  c("1", "2", "3", "4", "8"), labels = c("white", "black", "asian", "puerto rican", "other"))
           )
bw_data
```

*Part B:* Propose a regression model for birthweight.  
The model I created as 'fit1' is based on evidence supporting that mother's age and history of previous low birthweight births are predictive of additional low birthweight children.  I also included sex and presence of malformations that could impact weight so that we can control for these factors which may be potential confounders.
```{r}
bw_data = 
  bw_data %>% 
  mutate(
    malform = fct_infreq(malform),
    babysex = fct_infreq(babysex)
  )

fit1 = lm(bwt ~ momage + pnumlbw + babysex + malform, data = bw_data)
fit1

fit1 %>% 
  broom::tidy() %>%
  select(term, estimate, p.value) %>% 
  knitr::kable(digits = 3)


modelr::add_residuals(bw_data, fit1)
modelr::add_predictions(bw_data, fit1)

bw_data %>% 
  modelr::add_residuals(fit1) %>% 
  ggplot(aes(x = momage, y = resid)) + geom_point()

```

