---
title: "p8105_hw6_mk4022"
author: "Mary Knoop"
date: "11/23/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(modelr)
library(knitr)
set.seed(1)
```


**Question 1:**   
*Part A:*  Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).
```{r}
bw_data=
  read_csv ("./data/birthweight.csv") %>% 
  janitor::clean_names() %>%
  mutate(babysex = factor(babysex, levels =  c("1", "2"), labels = c("male", "female")),
         frace = factor(frace, levels =  c("1", "2", "3", "4", "8", "9"), labels = c("white", "black", "asian", "puerto rican", "other", "unknown")),
         malform = factor(malform, levels =  c("0", "1"), labels = c("absent", "present")),
         mrace = factor(mrace, levels =  c("1", "2", "3", "4", "8"), labels = c("white", "black", "asian", "puerto rican", "other"))
           )
bw_data
```

*Part B:* Propose a regression model for birthweight.  
* The model I created as 'fit1' is based on evidence supporting that mother's age, smoking status/quantity, and  history of previous low birthweight births are predictive of giving birth to a child with low birthweight. I also included sex and presence of malformations that could impact weight so that we can control for these factors which may be potential confounders.  
* I will use the A1C of each model to determine the best model.  
* The A1C of model 'fit1' is 66364.13.   
* The A1C of model 'fit2' is 62770.34.  
* The A1C of model 'fit3' is 61501.19.   
* According to the A1C's of the above three models, the best model for this dataset is model 'fit3' which has the lowest A1C. 
  

```{r}
##### Model 1 #####
#set references for factor variables
bw_data = 
  bw_data %>% 
  mutate(
    malform = fct_infreq(malform),
    babysex = fct_infreq(babysex)
  )

#create linear model
fit1 = lm(bwt ~ momage + pnumlbw + smoken + babysex + malform, data = bw_data)
fit1

#tidy output
fit1 %>% 
  broom::tidy() %>%
  select(term, estimate, p.value) %>% 
  knitr::kable(digits = 3)

#view A1C and other output parameters
fit1 %>% 
  broom::glance() 

#add residuals and predictions
modelr::add_residuals(bw_data, fit1)
modelr::add_predictions(bw_data, fit1)

#plot residuals by mom's age
bw_data %>% 
  modelr::add_residuals(fit1) %>% 
  ggplot(aes(x = momage, y = resid)) + geom_point()


##### Model 2 #####

#create linear model
fit2 = lm(bwt ~ blength + gaweeks, data = bw_data)
fit2

#tidy output
fit2 %>% 
  broom::tidy() %>%
  select(term, estimate, p.value) %>% 
  knitr::kable(digits = 3)

#view A1C and other output parameters
fit2 %>% 
  broom::glance() 

#add residuals and predictions
modelr::add_residuals(bw_data, fit2)
modelr::add_predictions(bw_data, fit2)

#plot residuals by mgestations age in weeks
bw_data %>% 
  modelr::add_residuals(fit2) %>% 
  ggplot(aes(x = gaweeks, y = resid)) + geom_point()



##### Model 3 #####

#set references for factor variables
bw_data = 
  bw_data %>% 
  mutate(
    babysex = fct_infreq(babysex)
  )

#create linear model
fit3 = lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data = bw_data)
fit3

#tidy output
fit3 %>% 
  broom::tidy() %>%
  select(term, estimate, p.value) %>% 
  knitr::kable(digits = 3)

#view A1C and other output parameters
fit3 %>% 
  broom::glance() 

#add residuals and predictions
modelr::add_residuals(bw_data, fit3)
modelr::add_predictions(bw_data, fit3)

#plot residuals by baby's sex
bw_data %>% 
  modelr::add_residuals(fit3) %>% 
  ggplot(aes(x = babysex, y = resid)) + geom_violin()

```

